Can Auxiliary Indicators Improve COVID-19 Forecasting and Hotspot Prediction?

### 2021-06-22 forecast/paper/supplement-text.tex
Go to 95% and fix up the leading/lagging

In Section 2.D of the main text, we discuss the extent to which the indicators
are leading or lagging case rates during different periods. To define the amount
of leadingness or laggingness, we use the cross correlation function (CCF)
between the two time series. The CCF is of an indicator $X_{\ell,t}$ and case
rates $Y_{\ell,t}$ is defined as
\[
\CCF_{\ell,t}(a) = \frac{1}{n_a} \sum_{i=1}^{n_a} \left( ( X_{\ell,t,i+a} -
  \overline{X}_{\ell,t}) / s^{X}_{\ell,t}\right) \left( ( Y_{\ell,t,i} -
  \overline{Y}_{\ell,t}) / s^{Y}_{\ell,t}\right), 
\]
where $n_a$ is the number of available time points when $X_{\ell,t}$ has been
shifted in time by $a$ days, and $\overline{X}_{\ell,t}$, $s^X_{\ell,t}$
are the sample mean and
standard deviation of $X_{\ell,t}$ (respectively $Y_{\ell,t}$). The result is a
sequence of Pearson correlations for each selected $a$, where we align
$X_{\ell,t}$ with the values of $Y_{\ell,t}$ that occurred $a$ days earlier.
Thus, for $a>0$, $\CCF_{\ell,t}(a)>0$ indicates that
$Y_{\ell,t}$ is moving together with the values of $X_{\ell,t}$ that occur $a$
days in the future. In this case we say that $X_{\ell,t}$ is lagging
$Y_{\ell,t}$. For $a<0$, $\CCF_{\ell,t}(a)>0$ means that $Y_{\ell,t}$ is
positively correlated with $X_{\ell,t}$ shifted earlier, so we say that
$X_{\ell,t}$ leads $Y_{\ell,t}$.

Figure \ref{fig:ccf-dv-finalized} shows the standardized signals for the HRR
containing Charlotte, North Carolina, from August 1, 2020 until the end of
September. These are the same signals shown in Figure 1 in the manuscript. To
define ``leadingness'' we compute $\CCF_{\ell,t}(a)$ for each $a\in
\{-15,\ldots,15\}$ using the 56 days leading up to time $t$ (this is the same
amount of data used to train the forecasters: 21 days of training data, 21 days
to get the response at $a=21$, and 14 days for the longest lagged value). The
orange dashed horizontal line represents the 97.5\% significance threshold for
correlations based on 56 observations. Any correlations larger in magnitude than
this value are considered statistically significant under the null hypothesis of
no relationship. We define leadingness to be the sum of the significant
correlations that are leading (those above the dashed line with $a<0$) while
laggingness is the same but for $a>0$. In this case, there are no significant
correlations, so both scores will be 0: on September 28, \dv~is neither
leading nor lagging cases.

>>>>

In Section 2.D of the main text, we discuss the extent to which the indicators
are leading or lagging case rates during different periods. To define the amount
of leadingness or laggingness, we use the cross correlation function (CCF)
between the two time series. The $\CCF_\ell(a)$ of an indicator $X_{\ell}$ and case
rates $Y_{\ell}$ is defined as
% \[
% \CCF_{\ell,t}(a) = \frac{1}{n_a} \sum_{i=1}^{n_a} \left( ( X_{\ell,t,i+a} -
%   \overline{X}_{\ell,t}) / s^{X}_{\ell,t}\right) \left( ( Y_{\ell,t,i} -
%   \overline{Y}_{\ell,t}) / s^{Y}_{\ell,t}\right), 
% \]
% where $n_a$ is the number of available time points when $X_{\ell,t}$ has been
% shifted in time by $a$ days, and $\overline{X}_{\ell,t}$, $s^X_{\ell,t}$
% are the sample mean and
% standard deviation of $X_{\ell,t}$ (respectively $Y_{\ell,t}$). The result is a
their Pearson correlation where $X_\ell$ has been
% sequence of Pearson correlations for each selected $a$, where we
aligned
with the values of $Y_{\ell}$ that occurred $a$ days earlier.
Thus, for any $a>0$, $\CCF_{\ell}(a)>0$ indicates that
$Y_{\ell,t}$ is moving together with the values of $X_{\ell,t+a}$.
% days in the future.
In this case we say that $X_{\ell}$ is lagging
$Y_{\ell}$. For $a<0$, $\CCF_{\ell}(a)>0$ means that $Y_{\ell,t}$ is
positively correlated with $X_{\ell,t-a}$, so we say that
$X_{\ell}$ leads $Y_{\ell}$.

Figure \ref{fig:ccf-dv-finalized} shows the standardized signals for the HRR
containing Charlotte, North Carolina, from August 1, 2020 until the end of
September. These are the same signals shown in Figure 1 in the manuscript but
using finalized data. To
define ``leadingness'' we compute $\CCF_{\ell}(a)$ for each $a\in
\{-15,\ldots,15\}$ using the 56 days leading up to the target date (this is the same
amount of data used to train the forecasters: 21 days of training data, 21 days
to get the response at $a=21$, and 14 days for the longest lagged value). The
orange dashed horizontal line represents the 95\% significance threshold for
correlations based on 56 observations. Any correlations larger in magnitude than
this value are considered statistically significant under the null hypothesis of
no relationship. We define leadingness to be the sum of the significant
correlations that are leading (those above the dashed line with $a<0$) while
laggingness is the same but for $a>0$. In this case, there are no significant
correlations, so both scores will be 0: on September 28, \dv~is neither
leading nor lagging cases.

### 2021-06-22 forecast/paper/supplement.Rmd
Go to 95% and fix up the leading/lagging

```{r lead-lag-processing}
# see lead-lag-analysis.R for signal downloading and processing
lead_lag <- function(x, y, lag.max = NULL, na.len = 14) {
  if (sum(is.na(x)) >= na.len || sum(is.na(y)) >= na.len) {
    return(c(leading = NA, lagging = NA))
  }
  out <- ccf(x, y, lag.max = lag.max, plot = FALSE, na.action = na.omit)
  clim <- qnorm((1 + .95) / 2) / sqrt(out$n.used)
  lag_vec <- drop(out$lag)
  cc <- drop(out$acf)
  cc <- cc * (cc > clim)
  return(c(leading = mean(cc[lag_vec < 0]),
           lagging = mean(cc[lag_vec > 0])))
}

>>>>

```{r lead-lag-processing}
# see lead-lag-analysis.R for signal downloading and processing
lead_lag <- function(x, y, lag.max = NULL, na.len = 14) {
  if (sum(is.na(x)) >= na.len || sum(is.na(y)) >= na.len) {
    return(c(leading = NA, lagging = NA))
  }
  out <- ccf(x, y, lag.max = lag.max, plot = FALSE, na.action = na.omit)
  clim <- qnorm((1 + .9) / 2) / sqrt(out$n.used)
  lag_vec <- drop(out$lag)
  cc <- drop(out$acf)
  cc <- cc * (cc > clim)
  return(c(leading = mean(cc[lag_vec < 0]),
           lagging = mean(cc[lag_vec > 0])))
}
